\LARGE Untangling the link between traits, size and growth rate in plants

### 2015-04-29 ms.tex
New conceptual figure #37
Download baad and build figure from data. Closes #26
General tidy, remove outdated material, rename variables

% Fancy HEADER
\usepackage{fancyhdr}
\pagestyle{fancy}
\pagenumbering{arabic}
\lhead{\itshape{\nouppercase{\leftmark}}}
\chead{}
\rhead{\thepage}
%\lfoot{v }
\cfoot{}
%\rfoot{\thepage}

@@

\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Untangling the link between traits, size and growth rate in plants},
            colorlinks=true,
            citecolor=black,
            urlcolor=grey,
            linkcolor=grey}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines

@@

The results of this new analysis
(Fig. \ref{f-BCI}) align closely with our theoretical predictions (Table
\ref{tab:predictions}) showing how the effect of traits on growth
changes with plant size, with different traits exhibiting distinct
size-dependent effects.

@@

Predictions
are compared against data from both a long-term
plot \citep{condit-2012} and field
experiment \citep{kitajima-2013} at Barro Colorado Island, Panama.
We use a new method to estimate species potential growth rate,
accounting for size specific effects (Fig. \ref{S-quantile_examples},
see Methods for details). This new method allowed us to extract a more
definite signal of trait effects than in previous
studies \citep{wright-2010}.

@@

We begin with a standard model for the amount of biomass available for
growth, $dP / \textrm{d}t$, given by the difference between income
(total photosynthesis) and losses (respiration and turnover) within the
plant \citep{makela-1997, Thornley-2000, falster-2011}:

\begin{equation}\label{eq:dPdt}
\underbrace{\strut \frac{dP}{\textrm{d}t}}_\textrm{net biomass production}
  = \underbrace{\strut Y}_\textrm{yield}
    \big( \underbrace{\strut a_\textrm{l} \, A}_\textrm{photosynthesis} -
     \underbrace{\strut \,\sum_\textrm{i=l,b,s,r}{m_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)
    - \underbrace{\strut \sum_\textrm{i=l,b,s,r}{m_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}.
\end{equation}

Here, $m,r$, and $k$ refer to the mass, respiration rate, and
turnover rate of different tissues, denoted by subscripts $l$=leaves,
$b$=bark, $s$=sapwood and $r$=roots. $A$ is the assimilation
rate of CO$_2$ per leaf area and $Y$ is yield: the fraction of
assimilated carbon fixed in biomass (the remaining fraction being lost
as growth respiration) (see Table \ref{tab:params} for units and
definitions). Photosynthesis is proportional to leaf area,
$a_\textrm{l} = m_\textrm{l} / \phi$, where $\phi$ is leaf mass per area.
The total mass of living tissues $m_\textrm{t}=m_\textrm{l}+m_\textrm{b}+m_\textrm{s}+m_\textrm{r}.$

The only potentially controversial element of in eq. \ref{eq:dPdt} is that it assumes tissues lost via turnover are replaced before energy is allocation to either new growth or reproduction \citep{Thornley-2000}. This assumption is likely to hold true for most woody plants, because the ..... However, the assumption may not hold for some herbs or annuals, where the swicth to reproduction often entails run-down in the vegtative part of the plant (refs XX) \citep{ Thornley-2000}.

@@

\begin{equation} \label{eq:dhdt}
\underbrace{\strut\frac{\textrm{d}h}{\textrm{d}t}}_\textrm{height growth rate}= \underbrace{\strut\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}}_\textrm{architecture}
\times \underbrace{\strut\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}}_\textrm{marginal leaf deployment}
\times \underbrace{\strut\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}}_\textrm{allocation to growth}
\times \underbrace{\strut\frac{\textrm{d}P}{\textrm{d}t}}_\textrm{mass production}.
\end{equation}

@@

The second term, $\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}$,
accounts for the marginal cost of deploying an additional unit of leaf
area, including construction of the leaf itself and various support
structures. As such, $\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}$
can itself be expressed as a sum of construction costs per unit leaf
area:

\begin{equation}\label{eq:daldmt}
\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}
= \big(\phi
 + \frac{\textrm{d}m_\textrm{s}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{b}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}\big)^{-1}.
\end{equation}

The third term in eq \ref{eq:dhdt},
$\textrm{d}m_\textrm{t} / \textrm{d}P$, gives the fraction of net
biomass production (eq. \ref{eq:dPdt}) that is allocated to growth
rather than reproduction or storage.

@@

\begin{equation}\label{eq:dast}
\frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}=
\left(\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}a_\textrm{l}}\right) \times
\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}} \times \frac{\textrm{d}m_\textrm{t}}{\textrm{d}P} \times \frac{\textrm{d}P}{\textrm{d}t} +\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t} .
\end{equation}

@@

By definition, the main effect of increasing $\phi$ is to decrease the
marginal leaf deployment per mass (eq. \ref{eq:daldmt}) and thus reduce
growth rate. However, to fully account for the influence of $\phi$ on
growth we must also account for the decrease in leaf turnover that
results from superior leaf construction \citep{wright_world-2004}.
This is achieved by linking the turnover rate of leaf in eq.
\ref{eq:dPdt} to $\phi$. We use an allometric scaling relationship of
the form $k_\textrm{l}=\alpha_4 \, \phi^{-\beta_4}$, which has been
shown to describe patterns across diverse
ecosystems \citep{wright_world-2004} (Fig. \ref{fS-leaf}).

@@

Unlike leaf and stem traits, height at maturation ($h_m$) moderates
growth by adjusting the amount of energy invested in growth, i.e.~the
term $\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}$ in eqs.
\ref{eq:dhdt} and \ref{eq:dast}.

@@

\subsection{Changes in growth rate with size}\label{changes-in-growth-rate-with-size}

This new growth model captures the intrinsically size-dependent nature
of plant growth (Table \ref{tab:predictions}a), recovering well-known
empirical patterns \citep{sillett-2010, king-2011}. Height growth shows a hump-shaped pattern
with size, first increasing then decreasing. This pattern results from
systematic changes in the four components of eq. \ref{eq:dhdt} with size
(Fig. \ref{f-hump}), including a strong decline in the fraction of plant
that is leaf declines with increasing size (Fig. \ref{f-mass_fraction}),
and increasing reproductive allocation. In contrast, basal-area growth
continues to increase with size \citep{sillett-2010,
@stephenson-2014}, due to an increasing influence of stem
turnover. Diameter growth shows a weakly hump shaped
curve \citep{herault-2011}, tapering off slightly at larger
sizes, in part because of allometric effects (eq. \ref{eq:dDdt}, but
mainly because of increased reproductive allocation in older trees
(Fig?). All growth measures decrease sharply with size when expressed as
relative growth rates \citep{iida-2014}.

\subsection{How traits influence growth rate}\label{how-traits-influence-growth-rate}

\subsubsection{LMA}\label{lma}

Unlike eq. \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and
\ref{eq:dDdt} predict a relationship between $\phi$ and growth that
changes with plant size (Figs. \ref{f-BCI}, \ref{f-lma_growth_size}).
Decreasing $\phi$ has two impacts on growth rate. First, lower
$\phi$ increases marginal leaf deployment
($\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}$ by economising on
construction costs. Second, lower $\phi$ decreases net production
($\textrm{d}P / \textrm{d}t$), due to increased leaf turnover. Whether
lower $\phi$ increases growth thus depends on the relative magnitude
of these two effects. When plants are small the effect on leaf
deployment rate is bigger and so decreasing $\phi$ increases growth
rate. When plants are large, the influence of $\phi$ on leaf
deployment rate is diminished, because the costs of building other
supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
\ref{f-mass_fraction}). The net result is that at larger sizes, low
$\phi$ is no longer advantageous for growth (Fig.
\ref{f-lma_growth_size}d).

\subsubsection{Wood density}\label{wood-density}

@@

\subsubsection{hmat}\label{hmax}

 Greater height at maturation ($h_m$)
can thus lead to a growth advantage by increasing
$\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}$ (Table
\ref{tab:predictions}). At smaller sizes, there is no differentiation
among species, because most species are focusing on growth (Fig.
\ref{f-BCI}). At larger sizes, individuals of some species are
allocating a majority of their surplus energy to growth, leaving a
strong signal in potential growth rate (Fig. \ref{f-BCI}).

\subsubsection{Responisveness to light}\label{hmax}

Two other important predictions arise directly from linking traits to
growth rate (Table \ref{tab:predictions}c). First, traits moderate the
responsiveness of growth to changes in light environment. This response
arises simply because species with higher potential growth rate have
greater potential plasticity. Recent analyses have shown that species
with low $rho$ increase growth more substantially with incraeses in
light \citep{ruger-2012}. This is exactly the behaviour
exhibited in our model (Fig. \ref{f-growth_light}), at least for small
to intermediate sizes. At larger sizes (D=0.2m) however, the advantages
of low $\phi$ are lost (Fig. \ref{f-growth_light}). This results also
agrees with empirical results showing that growth for species with low
$rho$ actually decreases as size
increases \citep{ruger-2012}. In contrast, growth rate tends
to increase in size for the species with highest
$\rho$ \citep{ruger-2012}.

@@

Current theory emphasises the relationship between $\phi$ and the relative growth in plant mass
\citep{lambers-1992, cornelissen-1996, wright_cross-2000,  enquist-2007},
$\textrm{d}P / \textrm{d}t / m_\textrm{t}$. This relationship can be derived from eq. \ref{eq:dPdt}  by ...
(see Appendix \ref{app:traits-RGR}).

@@

\subsection{Shade tolerance}\label{how-traits-influence-growth-rate}

@@

@@

\begin{table}[h]
\caption{Equations for an allometric growth model, based on functional balance assumption in {\textbf a}. }
\centering

{\footnotesize  % smaller font, double space
 \begin{doublespace}
  \begin{tabular}{p{2.5cm}p{3.5cm}p{5cm}p{4cm} }\\ \hline
  Variable & Function & Marginal cost & Growth rate\\ \hline
  \multicolumn{4}{l}{\textbf{a) Assumed relationship to leaf area}} \\ \\
  height &
    $h = \alpha_1 \, a_\textrm{l}^{\beta_1}$ &
    $\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} =  \beta_1 \alpha_1 a_\textrm{l}^{\beta_1-1}$ &
    $\frac{\textrm{d}h}{\textrm{d}t}  = \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
  sapwoood area &
    $a_\textrm{ss} = \theta \, a_\textrm{l}$ &
    $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d} a_\textrm{l}} = \theta$ &
    $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}t}  =\frac{\textrm{d}a_\textrm{ss}}{\textrm{d} a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
  bark area &
    $a_\textrm{sb} = b \, \theta \, a_\textrm{l}$ &
    $\frac{\textrm{d}a_\textrm{sb}}{\textrm{d} a_\textrm{l}} = b \, \theta$ &
    $\frac{\textrm{d}a_\textrm{sb}}{\textrm{d}t} = \frac{\textrm{d}a_\textrm{sb}}{\textrm{d} a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\  \\
  \multicolumn{4}{l}{\textbf{b) Derived equation for mass of tissue }} \\
  leaf mass &
    $m_\textrm{l} = \phi \, a_\textrm{l} $ &
    $\frac{\textrm{d}m_\textrm{l}}{\textrm{d}a_\textrm{l}} = \phi$ &
    $\frac{\textrm{d}m_\textrm{l}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{l}}{\textrm{d}a_\textrm{l}}  \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
  sapwood mass &
    $m_\textrm{ss} = \theta \, \rho \, \eta_c \, a_\textrm{l} \, h $ &
    $\frac{\textrm{d}m_\textrm{ss}}{\textrm{d}a_\textrm{l}} = \theta\, \rho\, \eta_c\, \big( h + a_\textrm{l}\, \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \big)$ &
    $\frac{\textrm{d}m_\textrm{ss}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{ss}}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
  bark mass &
    $m_\textrm{sb} = b\, \theta \, \rho \, \eta_c \, a_\textrm{l} \, h $ &
    $\frac{\textrm{d}m_\textrm{sb}}{\textrm{d}a_\textrm{l}} = b \, \theta \, \rho \, \eta_c\big( h + a_\textrm{l} \, \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \big)$ &
    $\frac{\textrm{d}m_\textrm{sb}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{sb}}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\

  root mass &
    $m_\textrm{r} = \alpha_3 \, a_\textrm{l}$ &
    $\frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}} = \alpha_3$  &
    $\frac{\textrm{d}m_\textrm{r}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}  \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\ \\

@@

\begin{table}[h]
\caption{Effect of traits and related trade-offs on demography}
\centering
{\footnotesize  % smaller font, double space
  \begin{doublespace}
  \begin{tabular}[c]{l|c|ccccc|c|c}
  \multicolumn{9}{c}{}\\ \hline
  & \multicolumn{8}{c}{\textbf{Effect of trait on}}\\
  & \textbf{Birth} & \multicolumn{5}{c|}{\textbf{Growth}} & \textbf{Mortality} & \textbf{Fecundity} \\
  & $h_0$
  & $\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}$
  & $\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}$
  & $\frac{\textrm{d}P}{\textrm{d}t}$
  & $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}}$
  & $\frac{1}{k_s}$ & &  \\\hline
  \textbf{This study}&&&&&&&&\\
  Leaf-construction cost, $\phi$ & & $\downarrow$ & $\uparrow$ & & & & \\
  Wood-construction cost, $\rho$ & & $\downarrow$ &  & & & $\downarrow$ & $\downarrow$  \\
  Height at maturation, $h_m$ & & &$\uparrow$ & & & & & $\downarrow$  \\
  &&&&&&&&\\\hline
  \textbf{Possible extensions}&&&&&&&&\\
  Sapwood area per leaf area, $\theta$ & & $\downarrow$& & $\uparrow$ & $\uparrow$ & & \\
  Leaf-nitrogen per area & & &$\downarrow$$\uparrow$ & & & & & \\
  Seed size & $\uparrow$ & & & & & & & $\downarrow$ \\ \hline
  \end{tabular}
  \end{doublespace}
}
\label{tab:trade-offs}
\end{table}

>>>>

% Fancy HEADER
\usepackage{fancyhdr}
\pagestyle{fancy}
\setlength{\headheight}{15pt}
\pagenumbering{arabic}
\lhead{\itshape{\nouppercase{\leftmark}}}
\chead{}
\rhead{\thepage}
%\lfoot{v }
\cfoot{}
%\rfoot{\thepage}

@@

\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Untangling the link between traits, size and growth rate in plants},
            colorlinks=true,
            citecolor=black,
            urlcolor=grey,
            linkcolor=grey}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines

@@

The results of this new analysis
(Fig. \ref{f-BCI}) align closely with our theoretical predictions  showing how the effect of traits on growth
changes with plant size, with different traits exhibiting distinct
size-dependent effects.

@@

Predictions
are compared against data from both a long-term
plot \citep{condit-2012} and field
experiment \citep{kitajima-2013} at Barro Colorado Island, Panama.
We use a new method to estimate species potential growth rate,
accounting for size specific effects (Fig. \ref{fS-quantile_examples},
see Methods for details). This new method allowed us to extract a more
definite signal of trait effects than in previous
studies \citep{wright-2010}.

@@

We begin with a standard model for the amount of biomass available for
growth, $\textrm{d}b / \textrm{d}t$, given by the difference between income
(total photosynthesis) and losses (respiration and turnover) within the
plant \citep{makela-1997, Thornley-2000, falster-2011}:

\begin{equation}\label{eq:dbdt}
\underbrace{\strut \frac{\textrm{d}b}{\textrm{d}t}}_\textrm{net biomass production}
  = \underbrace{\strut y}_\textrm{yield}
    \big( \underbrace{\strut a_\textrm{l} \, p}_\textrm{photosynthesis} -
     \underbrace{\strut \,\sum_\textrm{i=l,b,s,r}{m_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)
    - \underbrace{\strut \sum_\textrm{i=l,b,s,r}{m_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}.
\end{equation}

Here, $m,r$, and $k$ refer to the mass, respiration rate, and
turnover rate of different tissues, denoted by subscripts $l$=leaves,
$b$=bark, $s$=sapwood and $r$=roots. $A$ is the assimilation
rate of CO$_2$ per leaf area and $y$ is yield: the fraction of
assimilated carbon fixed in biomass (the remaining fraction being lost
as growth respiration) (see Table \ref{tab:params} for units and
definitions). Photosynthesis is proportional to leaf area,
$a_\textrm{l} = m_\textrm{l} / \phi$, where $\phi$ is leaf mass per area.
The total mass of living tissues $m_\textrm{a}=m_\textrm{l}+m_\textrm{b}+m_\textrm{s}+m_\textrm{r}.$

The only potentially controversial element of in eq. \ref{eq:dbdt} is that it assumes tissues lost via turnover are replaced before energy is allocation to either new growth or reproduction \citep{Thornley-2000}. This assumption is likely to hold true for most woody plants, because the ..... However, the assumption may not hold for some herbs or annuals, where the swicth to reproduction often entails run-down in the vegtative part of the plant (refs XX) \citep{ Thornley-2000}.

@@

\begin{equation} \label{eq:dhdt}
\underbrace{\strut\frac{\textrm{d}h}{\textrm{d}t}}_\textrm{height growth rate}= \underbrace{\strut\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}}_\textrm{architecture}
\times \underbrace{\strut\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{a}}}_\textrm{marginal leaf deployment}
\times \underbrace{\strut\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}}_\textrm{allocation to growth}
\times \underbrace{\strut\frac{\textrm{d}b}{\textrm{d}t}}_\textrm{mass production}.
\end{equation}

@@

The second term, $\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{a}$,
accounts for the marginal cost of deploying an additional unit of leaf
area, including construction of the leaf itself and various support
structures. As such, $\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{a}$
can itself be expressed as a sum of construction costs per unit leaf
area:

\begin{equation}\label{eq:daldmt}
\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{a}}
= \big(\phi
 + \frac{\textrm{d}m_\textrm{s}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{b}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}\big)^{-1}.
\end{equation}

The third term in eq \ref{eq:dhdt},
$\textrm{d}m_\textrm{a} / \textrm{d}P$, gives the fraction of net
biomass production (eq. \ref{eq:dbdt}) that is allocated to growth
rather than reproduction or storage.

@@

\begin{equation}\label{eq:dast}
\frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}=
\left(\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}a_\textrm{l}}\right) \times
\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{a}} \times \frac{\textrm{d}m_\textrm{a}}{\textrm{d}b} \times \frac{\textrm{d}b}{\textrm{d}t} +\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t} .
\end{equation}

@@

By definition, the main effect of increasing $\phi$ is to decrease the
marginal leaf deployment per mass (eq. \ref{eq:daldmt}) and thus reduce
growth rate. However, to fully account for the influence of $\phi$ on
growth we must also account for the decrease in leaf turnover that
results from superior leaf construction \citep{wright_world-2004}.
This is achieved by linking the turnover rate of leaf in eq.
\ref{eq:dbdt} to $\phi$. We use an allometric scaling relationship of
the form $k_\textrm{l}=\alpha_4 \, \phi^{-\beta_4}$, which has been
shown to describe patterns across diverse
ecosystems \citep{wright_world-2004} (Fig. \ref{fS-leaf}).

@@

Unlike leaf and stem traits, height at maturation ($h_m$) moderates
growth by adjusting the amount of energy invested in growth, i.e.~the
term $\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}$ in eqs.
\ref{eq:dhdt} and \ref{eq:dast}.

@@

\subsection{Changes in growth rate with size}

This new growth model captures the intrinsically size-dependent nature
of plant growth, recovering well-known
empirical patterns \citep{sillett-2010, king-2011}. Height growth shows a hump-shaped pattern
with size, first increasing then decreasing. This pattern results from
systematic changes in the four components of eq. \ref{eq:dhdt} with size
(Fig. \ref{f-conceptual}), including a strong decline in the fraction of plant
that is leaf declines with increasing size (Fig. \ref{f-mass_fraction}),
and increasing reproductive allocation. In contrast, basal-area growth
continues to increase with size \citep{sillett-2010, stephenson-2014},
due to an increasing influence of stem
turnover. Diameter growth shows a weakly hump shaped
curve \citep{herault-2011}, tapering off slightly at larger
sizes, in part because of allometric effects (eq. \ref{eq:dDdt}, but
mainly because of increased reproductive allocation in older trees
(Fig?). All growth measures decrease sharply with size when expressed as
relative growth rates \citep{iida-2014}.

\subsection{How traits influence growth rate}

\subsubsection{LMA}

Unlike eq. \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and
\ref{eq:dDdt} predict a relationship between $\phi$ and growth that
changes with plant size (Figs. \ref{f-BCI}, \ref{f-lma_growth_size}).
Decreasing $\phi$ has two impacts on growth rate. First, lower
$\phi$ increases marginal leaf deployment
($\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{a}$ by economising on
construction costs. Second, lower $\phi$ decreases net production
($\textrm{d}P / \textrm{d}t$), due to increased leaf turnover. Whether
lower $\phi$ increases growth thus depends on the relative magnitude
of these two effects. When plants are small the effect on leaf
deployment rate is bigger and so decreasing $\phi$ increases growth
rate. When plants are large, the influence of $\phi$ on leaf
deployment rate is diminished, because the costs of building other
supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
\ref{f-mass_fraction}). The net result is that at larger sizes, low
$\phi$ is no longer advantageous for growth (Fig.
\ref{f-lma_growth_size}d).

\subsubsection{Wood density}

@@

\subsubsection{hmat}

 Greater height at maturation ($h_m$)
can thus lead to a growth advantage by increasing
$\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}$ . At smaller sizes, there is no differentiation
among species, because most species are focusing on growth (Fig.
\ref{f-BCI}). At larger sizes, individuals of some species are
allocating a majority of their surplus energy to growth, leaving a
strong signal in potential growth rate (Fig. \ref{f-BCI}).

\subsubsection{Responisveness to light}

Two other important predictions arise directly from linking traits to
growth rate. First, traits moderate the
responsiveness of growth to changes in light environment. This response
arises simply because species with higher potential growth rate have
greater potential plasticity. Recent analyses have shown that species
with low $rho$ increase growth more substantially with incraeses in
light \citep{ruger-2012}. This is exactly the behaviour
exhibited in our model (Fig. \ref{f-growth_light}), at least for small
to intermediate sizes. At larger sizes (D=0.2m) however, the advantages
of low $\phi$ are lost (Fig. \ref{f-growth_light}). This results also
agrees with empirical results showing that growth for species with low
$rho$ actually decreases as size
increases \citep{ruger-2012}. In contrast, growth rate tends
to increase in size for the species with highest
$\rho$ \citep{ruger-2012}.

@@

Current theory emphasises the relationship between $\phi$ and the relative growth in plant mass
\citep{lambers-1992, cornelissen-1996, wright_cross-2000,  enquist-2007},
$\textrm{d}P / \textrm{d}t / m_\textrm{a}$. This relationship can be derived from eq. \ref{eq:dbdt}  by ...
(see Appendix \ref{app:traits-RGR}).

@@

\subsection{Shade tolerance}\label{shade-tolerance}

@@

\section{Acknowledgments}

Figure 2 uses image by Tracey Saxby, IAN Image Library http://ian.umces.edu/imagelibrary.

@@

\begin{table}[ht]
 \caption{Key variables
%  $a = l + ss + sb +  r$,
%  $m_{sh}$,
%  $m_t = m_a + m_{sh}$,
%  $st = stem total$
}
\centering

{\footnotesize  % smaller font, double space
 \begin{doublespace}
  \begin{tabular}{p{2cm}p{2cm}p{7cm}}
  \hline
  Symbol & Unit & Description \\
  \hline

  $b$   & kg  & biomass originating from parent plant\\
  $m_i$ & kg  & mass retained on plant in tissue $i$\\
  $a_i$ & m$^2$  & surface-area or cross-sectional-area of tissue $i$\\
  $h$   & m  & height of plant\\
  $D$   & m  & diameter of plant stem\\

  $y$ & kg kg$^{-1}$ & Yield; ratio of carbon fixed in mass per carbon assimilated \\
  $p$ & kg yr$^{-1}$ m$^{-2}$  & photosynthesic rate per unit area \\
  $r_i$ & kg yr$^{-1}$ kg$^{-1}$  & respiration per unit tissue mass \\
  $k_i$ & yr$^{-1}$ & turnover rate for tissue \\
  $A$ & kg yr$^{-1}$ & total plant photosynthesis \\
  $R$ & kg yr$^{-1}$ & total plant respiration \\
  $T$ & kg yr$^{-1}$ & total plant turnover \\
  \hline
  \end{tabular}
\end{doublespace}}
\label{tab:definitions}
\end{table}

\begin{table}[ht]
\caption{Equations for an allometric growth model, based on functional balance assumption in {\textbf a}. }
\centering

{\footnotesize  % smaller font, double space
 \begin{doublespace}
  \begin{tabular}{p{2.5cm}p{3.5cm}p{5cm}p{4cm} }\\ \hline
  Variable & Function & Marginal cost & Growth rate\\ \hline
  \multicolumn{4}{l}{\textbf{a) Assumed relationship to leaf area}} \\ \\
  height &
    $h = \alpha_1 \, a_\textrm{l}^{\beta_1}$ &
    $\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} =  \beta_1 \alpha_1 a_\textrm{l}^{\beta_1-1}$ &
    $\frac{\textrm{d}h}{\textrm{d}t}  = \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
  sapwoood area &
    $a_\textrm{ss} = \theta \, a_\textrm{l}$ &
    $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d} a_\textrm{l}} = \theta$ &
    $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}t}  =\frac{\textrm{d}a_\textrm{ss}}{\textrm{d} a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
  bark area &
    $a_\textrm{sb} = b \, \theta \, a_\textrm{l}$ &
    $\frac{\textrm{d}a_\textrm{sb}}{\textrm{d} a_\textrm{l}} = b \, \theta$ &
    $\frac{\textrm{d}a_\textrm{sb}}{\textrm{d}t} = \frac{\textrm{d}a_\textrm{sb}}{\textrm{d} a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\  \\
  \multicolumn{4}{l}{\textbf{b) Derived equation for mass of tissue }} \\
  leaf mass &
    $m_\textrm{l} = \phi \, a_\textrm{l} $ &
    $\frac{\textrm{d}m_\textrm{l}}{\textrm{d}a_\textrm{l}} = \phi$ &
    $\frac{\textrm{d}m_\textrm{l}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{l}}{\textrm{d}a_\textrm{l}}  \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
  sapwood mass &
    $m_\textrm{ss} = \theta \, \rho \, \eta_c \, a_\textrm{l} \, h $ &
    $\frac{\textrm{d}m_\textrm{ss}}{\textrm{d}a_\textrm{l}} = \theta\, \rho\, \eta_c\, \big( h + a_\textrm{l}\, \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \big)$ &
    $\frac{\textrm{d}m_\textrm{ss}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{ss}}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
  bark mass &
    $m_\textrm{sb} = b\, \theta \, \rho \, \eta_c \, a_\textrm{l} \, h $ &
    $\frac{\textrm{d}m_\textrm{sb}}{\textrm{d}a_\textrm{l}} = b \, \theta \, \rho \, \eta_c\big( h + a_\textrm{l} \, \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \big)$ &
    $\frac{\textrm{d}m_\textrm{sb}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{sb}}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
  root mass &
    $m_\textrm{r} = \alpha_3 \, a_\textrm{l}$ &
    $\frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}} = \alpha_3$  &
    $\frac{\textrm{d}m_\textrm{r}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}  \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\ \\

@@

\begin{table}[ht]
\caption{Effect of traits and related trade-offs on demography}
\centering
{\footnotesize  % smaller font, double space
  \begin{doublespace}
  \begin{tabular}[c]{l|c|ccccc|c|c}
  \multicolumn{9}{c}{}\\ \hline
  & \multicolumn{8}{c}{\textbf{Effect of trait on}}\\
  & \textbf{Birth} & \multicolumn{5}{c|}{\textbf{Growth}} & \textbf{Mortality} & \textbf{Fecundity} \\
  & $h_0$
  & $\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{a}}$
  & $\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}$
  & $\frac{\textrm{d}b}{\textrm{d}t}$
  & $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}}$
  & $\frac{1}{k_s}$ & &  \\\hline
  \textbf{This study}&&&&&&&&\\
  Leaf-construction cost, $\phi$ & & $\downarrow$ & $\uparrow$ & & & & \\
  Wood-construction cost, $\rho$ & & $\downarrow$ &  & & & $\downarrow$ & $\downarrow$  \\
  Height at maturation, $h_m$ & & &$\uparrow$ & & & & & $\downarrow$  \\
  &&&&&&&&\\\hline
  \textbf{Possible extensions}&&&&&&&&\\
  Sapwood area per leaf area, $\theta$ & & $\downarrow$& & $\uparrow$ & $\uparrow$ & & \\
  Leaf-nitrogen per area & & &$\downarrow$$\uparrow$ & & & & & \\
  Seed size & $\uparrow$ & & & & & & & $\downarrow$ \\ \hline
  \end{tabular}
  \end{doublespace}
}
\label{tab:trade-offs}
\end{table}
